The Mass Distribution of Stellar-Mass Black Holes

### 2010-10-30 paper.tex
Vicky's changes.

@@

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Using Markov Chain
  Monte Carlo calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their relative Bayesian evidence as supported by the
  measurements.  The mass distribution of the low-mass systems is best
  fit by a power-law, while the distribution of the complete sample is
  best fit by the exponential model.  We examine the existence of a
  ``gap'' between the most massive neutron stars and least massive
  black holes, finding that the best model (the power law) fitted to
  the low-mass systems only gives a minimum black-hole mass of 4.3
  $\Msun$ (90\% confidence), while the best model (the exponential)
  fitted to all 20 systems gives a minumum black-hole mass of 4.5
  $\Msun$ (90\% confidence).  We conclude that our sample of black
  hole masses provides strong evidence of a gap between the maximum
  neutron star mass and minimum black hole mass.
\end{abstract}

@@

The most massive stars probably end their lives with a supernova
explosion, becoming stellar-mass black holes.  The mass distribution
of such black holes can provide important clues to the end stages of
evolution of these stars.  In addition, the mass distribution of
stellar-mass black holes is an important input in calculations of
rates of gravitational wave emission events from coalescing neutron
star-black hole and black hole-black hole binaries in the LIGO
gravitational wave observatory \citep{Abadie2010}.

Observations of low-mass X-ray binaries in both the optical and X-ray
can provide a measurement of the mass of the compact object in these
systems.  There are now enough such measurements where the mass of the
compact object is large enough to classify it as a black hole to
attempt to statistically reconstruct the mass distribution of these
stellar-mass black holes.

The first study of the mass distribution of stellar-mass black holes,
in \citet{Bailyn1998}, examined a sample of seven transient X-ray
binaries thought to contain a black hole, concluding in a Bayesian
analysis that the mass function was strongly peaked around seven solar
masses.  \citet{Bailyn1998} found evidence of a ``gap'' between the
least massive black hole and the upper limit for neutron star masses
of 3 $\Msun$ \citep{Kalogera1996}.  Such a gap is puzzling in light of
theoretical studies that predict a continuous distribution of compact
object supernova remnant masses with a smooth transition from neutron
stars to black holes \citep{Fryer2001}.

A more recent study \citep{Ozel2010}, also in a Bayesian framework,
largely confirmed these findings, examining 16 low-mass X-ray binary
systems containing black holes and finding a strongly peaked
distribution at $7.8 \pm 1.2 \, \Msun$.  \citet{Ozel2010} use two
models for the mass function: a Gaussian and a decaying exponential
with a minimum ``turn-on'' mass (motivated by the analytical model of
the black-hole mass function in \citet{Fryer2001}).  \citet{Ozel2010}
do not provide confidence limits for the minimum black hole mass,
instead discussing only the maximum-posterior parameters for their
models.  Nevertheless, it appears that their analysis confirms the
existence of a mass gap.  \citet{Ozel2010} discuss possible selection
effects that could lead to the appearance of a mass gap, but conclude
these effects could not produce the observed gap, which they therefore
claim is a real property of the black hole mass distribution.

@@

We use a Bayesian Markov-Chain Monte Carlo analysis to explore various
models for the black hole mass function for both samples.  We include
both parameteric models, such as a Gaussian, and non-parameteric
models where the mass function is represented by histograms with
various numbers of bins.  (Our set of models includes those of
\citet{Ozel2010} and \citet{Bailyn1998}.)  After computing posterior
distributions for the model parameters, we use model selection
techniques (including a new technique for efficient reversible-jump
MCMC \citep{Farr2010}) to compare the evidence for the various models
from both samples.

We find, like \citet{Ozel2010} and \citet{Bailyn1998}, strong evidence
for a mass gap among the best models for both systems, with the best
model for the lower-mass systems giving a minimum black hole mass of
4.3 $\Msun$ (at 90\% confidence), and the best model for the combined
sample of lower- and high-mass systems giving a minimum black hole
mass of 4.5 $\Msun$ (at 90\% confidence).  Among the models with lower
evidence, most also have a mass gap.

The theoretical model from \citet{Fryer2001}, a decaying exponential,
considered for the low-mass systems in \citet{Ozel2010}, is strongly
disfavored by our model selection.  We find that the lower-mass
systems are best described by a power law, followed closely by a
Gaussian (which is the other model from \citet{Ozel2010}).  We find
that the theoretical model from \citet{Fryer2001} is, however, the
preferred model for the combined sample of lower- and high-mass
systems.  A model with two separate Gaussian peaks also has high
evidence for the combined sample of systems.  While the failure of the
theoretically-motivated model to explain the lower-mass, Roche
lobe-filling systems suggests gaps in our understanding of the
formation of these systems, it is encouraging for population synthesis
studies \fixme{Vicky, do you have a couple of go-to papers for
  pop. synth we could cite here?} that the model of \citet{Fryer2001}
is a good fit to the full sample including both lower- and high-mass
systems.

@@

The 20 X-ray binary systems on which this study is based are listed in
Table \ref{tab:sources}.  We separate the systems into 15 low-mass
systems in which the central black hole appears to be fed by
roche-lobe overflow from the secondary, and 5 higher-mass systems in
which the black hole is fed via winds (these systems all have a
secondary that appears to be more massive than the black hole).  The
low- and higher-mass systems undoubtedly have different evolutionary
tracks, and therefore it is reasonable that they would have different
black-hole mass distributions.  We will first analyze the 15 low-mass
systems alone (Section \ref{sec:results-lower-mass}), and then the
complete sample of 20 systems (Section \ref{sec:higher-mass}).

@@

  \caption{\label{fig:low-masses} The mass distributions implied by
    Equation \eqref{eq:mass-function} and the assumed distributions on
    observational parameters given in Table \ref{tab:sources} for the
    lower-mass sources.  The significant asymmetry and long tails in
    many of these distributions are the result of the non-linear
    relationship (Equation \eqref{eq:mass-function}) between $M$, $f$,
    $q$, and $i$.}
\end{figure}

\begin{figure}
  \begin{center}
    \plotone{plots/high-masses}
  \end{center}
  \caption{\label{fig:high-masses} Mass distributions for the
    wind-fed, higher-mass systems computed from the distributions on
    observed data in Table \ref{tab:sources} using Equation
    \eqref{eq:mass-function}.  The asymmetry and long tails in these
    distributions are the result of the non-linear relationship
    between $M$, $f$, $q$, and $i$.}
\end{figure}

@@

In this section we describe the statistical analysis we will apply to
various models for the underlying mass distribution from which the
low-mass systems in Table \ref{tab:sources} were drawn.  Results of
the analysis appear in Section \ref{sec:results}.

@@

For a mass distribution with several parameters, $p(\vtheta | d)$
lives in a multi-dimensional space.  Previous works
\citep{Ozel2010,Bailyn1998} have considered models with only two
parameters, evaluating $p(\vtheta|d)$ on a grid.  Many of our models
for the underlying mass distribution have three or more parameters.
Exploring the entirety of these parameter spaces with a grid rapidly
becomes prohibitive as the number of parameters increases.  A more
efficient way to explore the distribution $p(\vtheta | d)$ is to use a
Markov Chain Monte Carlo (MCMC) method.  MCMC methods produce a chain
(sequence) of parameter samples, $\{ \vtheta_i \, | \, i = 1, \ldots
\}$, such that a particular parameter sample, $\vtheta$, appears in
the sequence with a frequency equal to its posterior probability,
$p(\vtheta|d)$.  In this way, regions of parameter space where
$p(\vtheta|d)$ is large are sampled densely while regions where
$p(\vtheta|d)$ is small are effectively ignored.

@@

We consider histograms with up to five bins in this work.  We will see
that the evidence for the histogram models (see Sections
\ref{sec:model-selection}, \ref{sec:lower-mass-model-selection}, and
\ref{sec:high-mass-model-selection}) from both the lower-mass and
complete datasets is decreasing as the number of bins reaches five,
indicating that increasing the number of bins beyond five would not
sufficiently improve the fit to the mass distribution to compensate
for the extra parameter-space volume implied by the additional
parameters.

@@

In this section we discuss the results of our MCMC analysis of the
posterior distributions of parameters for the models in Sections
\ref{sec:parametric-models} and \ref{sec:non-parametric-models}.  We
also discuss model selection results.  The results in Section
\ref{sec:results-lower-mass} apply to the lower-mass sample of
systems, while those of Section \ref{sec:higher-mass} apply to the
complete sample of systems.

\subsection{Lower-Mass Systems}
\label{sec:results-lower-mass}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/power-law}
  \end{center}
  \caption{\label{fig:power-law} Histograms of the marginalized
    distribution for the three parameters $\Mmin$, $\Mmax$, and
    $\alpha$ from the power-law model.  The marginalized distribution
    for $\alpha$ is broad, with $-11.8 < \alpha < 6.8$ enclosing 90\%
    of the probability.  We have $p(\alpha < 0) = 0.6$; the median
    value is $\alpha = -3.35$.  The broad distribution for $\alpha$
    (and the other parameters) is due to correlations between the
    parameters discussed in the main text; see Figure
    \ref{fig:power-law-2D}.}
\end{figure}

The marginalized distribution for $\alpha$ is broad, with
\begin{equation}
  -11.8 < \alpha < 6.8
\end{equation}
enclosing 90\% of the probability (excluding 5\% on each side).  We
have $p(\alpha < 0) = 0.6$.  The median value is $\alpha = -3.35$.
The broadness of the marginalized distribution for $\alpha$ comes from
the need to match the relatively narrow range in mass of the
lower-mass systems.  When $\alpha$ is negative, the resulting mass
distribution slopes down; $\Mmin$ is constrained to be near the lowest
mass of the observed black holes, while $\Mmax$ is essentially
irrelevant.  Conversely, when $\alpha$ is positive and the mass
distribution slopes up, $\Mmax$ must be close to the largest mass
observed, while $\Mmin$ is essentially irrelevant.  Figure
\ref{fig:power-law-2D} illustrates this effect, showing the
correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
When we include the higher-mass systems in the analysis, the long tail
will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
play for all values of $\alpha$.

@@

Figure \ref{fig:two-gaussian} shows the marginalized distributions for
the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
> 0.8$ with 62\% probability, clearly favoring the first Gaussian.
The distributions for $\mu_1$ and $\sigma_1$ are similar to those of
the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
that the first Gaussian is centered around the peaks of the lower-mass
distributions.  The second Gaussian's parameter distributions are much
broader.  The second Gaussian appears to be sampling the tail of the
mass samples.  In spite of the ability to match a more complicated
distribution, we find that this model is strongly disfavored relative
to the single-Gaussian model for this dataset:
$p(\textnormal{Gaussian}|d) / p(\textnormal{Two Gaussian}|d) \simeq
4.7$ (see Sections \ref{sec:model-selection} and
\ref{sec:lower-mass-model-selection} for discussion).

\begin{figure}
  \begin{center}
    \plotone{plots/two-gaussian}
  \end{center}
  \caption{\label{fig:two-gaussian} The marginal distributions for the
    five parameters of the two-Gaussian model.  The top panel is
    $\mu_1$ (solid histogram) and $\sigma_1$ (dashed histogram), the
    middle panel is $\mu_2$ (solid histogram) and $\sigma_2$ (dashed
    histogram), and the bottom panel is $\alpha$. We have $\alpha >
    0.8$ with 62\% probability, favoring the first of the two
    Gaussians.  The distributions for $\mu_1$ and $\sigma_1$ are
    similar to those of the single Gaussian model displayed in Figure
    \ref{fig:gaussian}; the second Gaussian's parameter distributions
    are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
    second Gaussian is attempting to fit the tail of the mass samples.
    The extra degrees of freedom in the distribution from the second
    Gaussian do not provide enough extra fitting power to compensate
    for the increase in parmeter space, however: the two-Gaussian
    model is disfavored relative to the single Gaussian by a factor of
    $4.7$ on this dataset (see Sections \ref{sec:model-selection} and
    \ref{sec:lower-mass-model-selection} for discussion).}
\end{figure}

@@

As the number of bins increases, the models are better able to capture
features of the mass distribution, but we find that the one-bin
histogram is the most probable of the histogram models for the
low-mass data (see Section \ref{sec:lower-mass-model-selection} for
discussion).  This occurs because the extra fitting power does not
sufficiently improve the fit to compensate for the vastly larger
parameter space of the models with more bins.

\subsubsection{Lower-Mass Model Selection}
\label{sec:lower-mass-model-selection}
We have performed a suite of 500 independent reversible-jump MCMCs
jumping between all the models (both parametric and non-parametric) in
Section \ref{sec:models} using the single-model MCMC samples to
construct an efficient jump proposal for each model as described above
(see Appendix \ref{sec:reversible-jump-mcmc}).  The numbers of counts in
each model are consistent across the MCMCs in the suite; Figure
\ref{fig:rj} displays the average probability for each model across
the suite, along with the 1-$\sigma$ errors on the average inferred
from the standard deviation of the model counts across the suite.
Table \ref{tab:rj} gives the numerical values of the average
probability for each model across the suite of MCMCs.

@@

\begin{table}
  \begin{center}
    \begin{tabular}{|l|r|}
      \hline
      Model & Relative Evidence \\
      \hline \hline
      Power Law (Section \ref{sec:power-law}) & 0.331488 \\
      \hline
      Gaussian (Section \ref{sec:gaussian}) & 0.288129 \\ 
      \hline
      Log Normal (Section \ref{sec:log-normal}) & 0.138435 \\
      \hline
      Exponential (Section \ref{sec:exponential}) & 0.0916218 \\
      \hline
      Two Gaussian (Section \ref{sec:gaussian}) & 0.0662577 \\
      \hline
      Histogram (1 Bin, Section \ref{sec:non-parametric-models}) &
      0.0641941 \\
      \hline
      Histogram (2 Bin, Section \ref{sec:non-parametric-models}) &
      0.015184 \\
      \hline 
      Histogram (3 Bin, Section \ref{sec:non-parametric-models}) &
      0.00332933 \\
      \hline
      Histogram (4 Bin, Section \ref{sec:non-parametric-models}) &
      0.000999976 \\
      \hline 
        Histogram (5 Bin, Section \ref{sec:non-parametric-models}) &
      0.0003614  \\
      \hline      
    \end{tabular}
  \end{center}
  \caption{\label{tab:rj} Relative probabilities of the various models
    from Section \ref{sec:models} implied by the lower-mass data.  These probabilities have been 
    computed from reversible-jump
    MCMC samples using the efficient jump proposal algorithm in Appendix \ref{sec:reversible-jump-mcmc}.}
\end{table}

@@

Figure \ref{fig:power-law-high} presents the marginalized
distribution for the three power-law parameters $\Mmin$, $\Mmax$, and
$\alpha$ (Section \ref{sec:power-law}) from an analysis including the
higher-mass systems.  The distribution for $\Mmax$ is quite broad
because the best fit power laws slope downward ($\alpha < 0$), making
this parameter less relevant.  The range $-5.05 \leq \alpha \leq
-1.77$ encloses 90\% of the probability; the median value of $\alpha$
is -3.23.  The presence of the higher-mass samples in the analysis
produces a distinctive tail, eliminating the correlations discussed in
Section \ref{sec:power-law} and displayed in Figure
\ref{fig:power-law-2D} for the lower-mass subset of the observations.

\begin{figure}
  \begin{center}
    \plotone{plots/power-law-high}
  \end{center}
  \caption{\label{fig:power-law-high} Histograms of the marginalized
    distribution for the three parameters $\Mmin$, $\Mmax$, and
    $\alpha$ from the power-law model including the higher-mass
    samples in the MCMC.  The distribution for $\Mmax$ is quite broad
    because the best fit power laws slope downward ($\alpha < 0$),
    making this parameter less relevant.  The range $-5.05 \leq \alpha
    \leq -1.77$ encloses 90\% of the probability; the median value of
    $\alpha$ is -3.23.  The presence of the higher-mass samples in the
    analysis produces a distinctive tail, eliminating the correlations
    discussed in Section \ref{sec:power-law} and displayed in Figure
    \ref{fig:power-law-2D} for the lower-mass subset of the
    observations. }
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/exp-cutoff-high}
  \end{center}
  \caption{\label{fig:exp-cutoff-high} The marginalized distributions
    for the exponential parameters $\Mmin$ and $M_0$ (Section
    \ref{sec:exponential}) from an analysis including the higher-mass
    systems.  The distribution for the scale mass, $M_0$, has moved to
    higher masses relative to Figure \ref{fig:exp-marginal} to fit the
    tail of the mass distribution; we now have $2.8292 \leq M_0 \leq
    7.9298$ with 90\% confidence, with median 4.7003.  The
    distribution for $\Mmin$ is less affected, though it has broadened
    somewhat toward low masses.}
\end{figure}

>>>>

\newcommand{\changed}[1]{{\bf #1 }}

@@

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Using Markov Chain
  Monte Carlo calculations, we model the mass distribution both
  parametrically---as a power law, exponential, gaussian, combination
  of two gaussians, or log-normal distribution---and
  non-parametrically---as histograms with varying numbers of bins.  We
  provide confidence bounds on the shape of the mass distribution in
  the context of each model and compare the models with each other by
  calculating their relative Bayesian evidence as supported by the
  measurements, \changed{taking into account the number of degrees of
    freedom of each model.}  The mass distribution of the low-mass
  systems is best fit by a power-law, while the distribution of the
  complete sample is best fit by the exponential model.  \changed{This
    difference clearly indicates that the character of the low mass
    X-ray binary systems differs from the higher-mass systems.}  We
  examine the existence of a ``gap'' between the most massive neutron
  stars and least massive black holes, finding that the best model
  (the power law) fitted to the low-mass systems only gives a minimum
  black-hole mass of 4.3 $\Msun$ (90\% confidence), while the best
  model (the exponential) fitted to all 20 systems gives a minumum
  black-hole mass of 4.5 $\Msun$ (90\% confidence).  We conclude that
  our sample of black hole masses provides strong evidence of a gap
  between the maximum neutron star mass and minimum black hole mass.
  \changed{Our results on the low-mass sample are in qualitative
    agreement with those of \citet{Ozel2010}, although our broad model
    selection analysis more reliably reveals the best-fit quantitative
    description of the underlying mass distribution.  The results on
    the complete sample of low- and higher-mass systems are in
    qualitative agreement with \citet{Fryer2001} although the presence
    of a mass gap remains theoretically unexplained.}
\end{abstract}

@@

The most massive stars probably end their lives with a supernova
explosion or a quiet core collapse, becoming stellar-mass black holes.
The mass distribution of such black holes can provide important clues
to the end stages of evolution of these stars.  In addition, the mass
distribution of stellar-mass black holes is an important input in
calculations of rates of gravitational wave emission events from
coalescing neutron star-black hole and black hole-black hole binaries
in the LIGO gravitational wave observatory \citep{Abadie2010}.

Observations of low-mass X-ray binaries in both the optical and X-ray
can provide a measurement of the mass of the compact object in these
systems.  \changed{The current sample of dynamically measured stellar
  mass black holes includes 15 low-mass, Roche lobe overflowing
  systems and 5 wind-fed higher-mass systems.  Hence, sophisticaed
  statistical analyses are possible.}

The first study of the mass distribution of stellar-mass black holes,
in \citet{Bailyn1998}, examined a sample of seven low-mass X-ray
binaries thought to contain a black hole, concluding in a Bayesian
analysis that the mass function was strongly peaked around seven solar
masses.  \citet{Bailyn1998} found evidence of a ``gap'' between the
least massive black hole and a ``safe'' upper limit for neutron star
masses of 3 $\Msun$ (e.g.\ \citet{Kalogera1996}).  Such a gap is
puzzling in light of theoretical studies that predict a continuous
distribution of compact object supernova remnant masses with a smooth
transition from neutron stars to black holes \citep{Fryer2001}.
\changed{(It should be noted that \citet{Fryer2001} considered binary
  evolution effects only heuristically and put forward some possible
  explanations for the gap from \citet{Bailyn1998} both in the context
  of selection effects or in connection to the energetics of supernova
  explosions.)}

\changed{Towards the end of our anlysis work, we became aware of a
  more recent study \citep{Ozel2010}, also in a Bayesian framework,
  analyzing the low-mass X-ray binary sample.}  It largely confirmed
these findings, examining 16 low-mass X-ray binary systems containing
black holes and finding a strongly peaked distribution at $7.8 \pm 1.2
\, \Msun$.  \citet{Ozel2010} use two models for the mass function: a
Gaussian and a decaying exponential with a minimum ``turn-on'' mass
(motivated by the analytical model of the black-hole mass function in
\citet{Fryer2001}).  \changed{\citet{Ozel2010} do not provide
  confidence limits for the minimum black hole mass, instead
  discussing only the model parameters at the peak of their posterior
  distributions.  They also do not perform any model selection
  analysis; thus, they give the distribution of parameters within each
  of their models, but cannot say which model is more likely to
  correspond to the true distribution of black hole masses.}
Nevertheless, it appears that their analysis confirms the existence of
a mass gap.  \citet{Ozel2010} discuss possible selection effects that
could lead to the appearance of a mass gap, but conclude these effects
could not produce the observed gap, which they therefore claim is a
real property of the black hole mass distribution.

@@

We use a Bayesian Markov-Chain Monte Carlo (MCMC) analysis to
quantitatively assess a wide range of models for the black hole mass
function for both samples.  We include both parameteric models, such
as a Gaussian, and non-parameteric models where the mass function is
represented by histograms with various numbers of bins.  (Our set of
models includes those of \citet{Ozel2010} and \citet{Bailyn1998}.)
After computing posterior distributions for the model parameters, we
use model selection techniques (including a new technique for
efficient reversible-jump MCMC \citep{Farr2010}) to compare the
evidence for the various models from both samples.

In agreement with \citet{Ozel2010} and \citet{Bailyn1998}, we find
strong evidence for a mass gap among the best models for both samples,
with the best model for the low-mass systems giving a minimum black
hole mass of 4.3 $\Msun$ (at 90\% confidence), and the best model for
the combined sample of lower- and high-mass systems giving a minimum
black hole mass of 4.5 $\Msun$ (at 90\% confidence).  Further, among
the models with lower evidence, most also have a mass gap.

\changed{We find that, for the low-mass X-ray binary sample, the
  theoretical model from \citet{Fryer2001}, a decaying exponential is
  strongly disfavored by our model selection.  We find that the
  low-mass systems are best described by a power law, followed
  closely by a Gaussian (which is the other model from
  \citet{Ozel2010}).  On the other hand, we find that the theoretical
  model from \citet{Fryer2001} is the preferred model for the combined
  sample of low- and high-mass X-ray binaries.  A model with two
  separate Gaussian peaks also has relatively high evidence for the
  combined sample of systems.  The difference in best-fitting model
  clearly indicates that the character of the low mass X-ray binary
  systems differs from the higher-mass systems.}

@@

The 20 X-ray binary systems on which this study is based are listed in
Table \ref{tab:sources}.  We separate the systems into 15 low-mass
systems in which the central black hole appears to be fed by
roche-lobe overflow from the secondary, and 5 higher-mass systems in
which the black hole is fed via winds (these systems all have a
secondary that appears to be more massive than the black hole).  The
low- and higher-mass systems undoubtedly have different evolutionary
tracks, and therefore it is reasonable that they would have different
black-hole mass distributions.  We will first analyze the 15 low-mass
systems alone (Section \ref{sec:results-low-mass}), and then the
complete sample of 20 systems (Section \ref{sec:higher-mass}).

@@

  \caption{\label{fig:low-masses} The individual mass distributions
    implied by Equation \eqref{eq:mass-function} and the assumed
    distributions on observational parameters $f$, $q$, and $i$ given
    in Table \ref{tab:sources} for the low-mass sources.  The
    significant asymmetry and long tails in many of these
    distributions are the result of the non-linear relationship
    (Equation \eqref{eq:mass-function}) between $M$, $f$, $q$, and
    $i$.}
\end{figure}

\begin{figure}
  \begin{center}
    \plotone{plots/high-masses}
  \end{center}
  \caption{\label{fig:high-masses} Mass distributions for the
    wind-fed, higher-mass systems computed from the distributions on
    observed data in Table \ref{tab:sources} using Equation
    \eqref{eq:mass-function}.  (Similar to Figure
    \ref{fig:low-masses}.)  The asymmetry and long tails in these
    distributions are the result of the non-linear relationship
    between $M$, $f$, $q$, and $i$.}
\end{figure}

@@

In this section we describe the statistical analysis we will apply to
various models for the underlying mass distribution from which the
low-mass sample and the complete sample of X-ray binary systems in
Table \ref{tab:sources} were drawn.  The results of our analysis are
presented in Section \ref{sec:results}.

@@

For a mass distribution with several parameters, $p(\vtheta | d)$
lives in a multi-dimensional space.  Previous works
\citep{Ozel2010,Bailyn1998} have considered models with only two
parameters; for such models evaluating $p(\vtheta|d)$ on a grid may be
a reliable method.  Many of our models for the underlying mass
distribution have three or more parameters.  Exploring the entirety of
these parameter spaces with a grid rapidly becomes prohibitive as the
number of parameters increases.  A more efficient way to explore the
distribution $p(\vtheta | d)$ is to use a Markov Chain Monte Carlo
(MCMC) method.  MCMC methods produce a chain (sequence) of parameter
samples, $\{ \vtheta_i \, | \, i = 1, \ldots \}$, such that a
particular parameter sample, $\vtheta$, appears in the sequence with a
frequency equal to its posterior probability, $p(\vtheta|d)$.  In this
way, regions of parameter space where $p(\vtheta|d)$ is large are
sampled densely while regions where $p(\vtheta|d)$ is small are
effectively ignored.

@@

We consider histograms with up to five bins in this work.  We will see
that the evidence for the histogram models (see Sections
\ref{sec:model-selection}, \ref{sec:low-mass-model-selection}, and
\ref{sec:high-mass-model-selection}) from both the low-mass and
complete datasets is decreasing as the number of bins reaches five,
indicating that increasing the number of bins beyond five would not
sufficiently improve the fit to the mass distribution to compensate
for the extra parameter-space volume implied by the additional
parameters.

@@

In this section we discuss the results of our MCMC analysis of the
posterior distributions of parameters for the models in Sections
\ref{sec:parametric-models} and \ref{sec:non-parametric-models}.  We
also discuss model selection results.  The results in Section
\ref{sec:results-low-mass} apply to the low-mass sample of
systems, while those of Section \ref{sec:higher-mass} apply to the
complete sample of systems.

\subsection{Low-Mass Systems}
\label{sec:results-low-mass}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/power-law}
  \end{center}
  \caption{\label{fig:power-law} Histograms of the marginalized
    distribution for the three parameters $\Mmin$ (top, left), $\Mmax$
    (top, right), and $\alpha$ (bottom) from the power-law model.  The
    marginalized distribution for $\alpha$ is broad, with $-11.8 <
    \alpha < 6.8$ enclosing 90\% of the probability.  We have
    $p(\alpha < 0) = 0.6$; the median value is $\alpha = -3.35$.  The
    broad distribution for $\alpha$ (and the other parameters) is due
    to correlations between the parameters discussed in the main text;
    see Figure \ref{fig:power-law-2D}.}
\end{figure}

The marginalized distribution for $\alpha$ is broad, with
\begin{equation}
  -11.8 < \alpha < 6.8
\end{equation}
enclosing 90\% of the probability (excluding 5\% on each side).  We
have $p(\alpha < 0) = 0.6$.  The median value is $\alpha = -3.35$.
The broadness of the marginalized distribution for $\alpha$ comes from
the need to match the relatively narrow range in mass of the
low-mass systems.  When $\alpha$ is negative, the resulting mass
distribution slopes down; $\Mmin$ is constrained to be near the lowest
mass of the observed black holes, while $\Mmax$ is essentially
irrelevant.  Conversely, when $\alpha$ is positive and the mass
distribution slopes up, $\Mmax$ must be close to the largest mass
observed, while $\Mmin$ is essentially irrelevant.  Figure
\ref{fig:power-law-2D} illustrates this effect, showing the
correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
When we include the higher-mass systems in the analysis, the long tail
will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
play for all values of $\alpha$.

@@

Figure \ref{fig:two-gaussian} shows the marginalized distributions for
the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
> 0.8$ with 62\% probability, clearly favoring the first Gaussian.
The distributions for $\mu_1$ and $\sigma_1$ are similar to those of
the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
that the first Gaussian is centered around the peaks of the low-mass
distributions.  The second Gaussian's parameter distributions are much
broader.  The second Gaussian appears to be sampling the tail of the
mass samples.  In spite of the ability to match a more complicated
distribution, we find that this model is strongly disfavored relative
to the single-Gaussian model for this dataset:
$p(\textnormal{Gaussian}|d) / p(\textnormal{Two Gaussian}|d) \simeq
4.7$ (see Sections \ref{sec:model-selection} and
\ref{sec:low-mass-model-selection} for discussion).

\begin{figure}
  \begin{center}
    \plotone{plots/two-gaussian}
  \end{center}
  \caption{\label{fig:two-gaussian} The marginal distributions for the
    five parameters of the two-Gaussian model.  The top panel is
    $\mu_1$ (solid histogram) and $\sigma_1$ (dashed histogram), the
    middle panel is $\mu_2$ (solid histogram) and $\sigma_2$ (dashed
    histogram), and the bottom panel is $\alpha$. We have $\alpha >
    0.8$ with 62\% probability, favoring the first of the two
    Gaussians.  The distributions for $\mu_1$ and $\sigma_1$ are
    similar to those of the single Gaussian model displayed in Figure
    \ref{fig:gaussian}; the second Gaussian's parameter distributions
    are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
    second Gaussian is attempting to fit the tail of the mass samples.
    The extra degrees of freedom in the distribution from the second
    Gaussian do not provide enough extra fitting power to compensate
    for the increase in parmeter space, however: the two-Gaussian
    model is disfavored relative to the single Gaussian by a factor of
    $4.7$ on this dataset (see Sections \ref{sec:model-selection} and
    \ref{sec:low-mass-model-selection} for discussion).}
\end{figure}

@@

As the number of bins increases, the models are better able to capture
features of the mass distribution, but we find that the one-bin
histogram is the most probable of the histogram models for the
low-mass data (see Section \ref{sec:low-mass-model-selection} for
discussion).  This occurs because the extra fitting power does not
sufficiently improve the fit to compensate for the vastly larger
parameter space of the models with more bins.

\subsubsection{Low-Mass Model Selection}
\label{sec:low-mass-model-selection}
We have performed a suite of 500 independent reversible-jump MCMCs
jumping between all the models (both parametric and non-parametric)
described in Section \ref{sec:models} using the single-model MCMC
samples to construct an efficient jump proposal for each model as
described above (see Appendix \ref{sec:reversible-jump-mcmc}).  The
numbers of counts in each model are consistent across the MCMCs in the
suite; Figure \ref{fig:rj} displays the average probability for each
model across the suite, along with the 1-$\sigma$ errors on the
average inferred from the standard deviation of the model counts
across the suite.  Table \ref{tab:rj} gives the numerical values of
the average probability for each model across the suite of MCMCs.

@@

\begin{table}
  \begin{center}
    \begin{tabular}{|l|r|}
      \hline
      Model & Relative Evidence \\
      \hline \hline
      Power Law (Section \ref{sec:power-law}) & 0.331488 \\
      \hline
      Gaussian (Section \ref{sec:gaussian}) & 0.288129 \\ 
      \hline
      Log Normal (Section \ref{sec:log-normal}) & 0.138435 \\
      \hline
      Exponential (Section \ref{sec:exponential}) & 0.0916218 \\
      \hline
      Two Gaussian (Section \ref{sec:gaussian}) & 0.0662577 \\
      \hline
      Histogram (1 Bin, Section \ref{sec:non-parametric-models}) &
      0.0641941 \\
      \hline
      Histogram (2 Bin, Section \ref{sec:non-parametric-models}) &
      0.015184 \\
      \hline 
      Histogram (3 Bin, Section \ref{sec:non-parametric-models}) &
      0.00332933 \\
      \hline
      Histogram (4 Bin, Section \ref{sec:non-parametric-models}) &
      0.000999976 \\
      \hline 
        Histogram (5 Bin, Section \ref{sec:non-parametric-models}) &
      0.0003614  \\
      \hline      
    \end{tabular}
  \end{center}
  \caption{\label{tab:rj} Relative probabilities of the various models
    from Section \ref{sec:models} implied by the low-mass data.  These probabilities have been 
    computed from reversible-jump
    MCMC samples using the efficient jump proposal algorithm in Appendix \ref{sec:reversible-jump-mcmc}.}
\end{table}

@@

Figure \ref{fig:power-law-high} presents the marginalized
distribution for the three power-law parameters $\Mmin$, $\Mmax$, and
$\alpha$ (Section \ref{sec:power-law}) from an analysis including the
higher-mass systems.  The distribution for $\Mmax$ is quite broad
because the best fit power laws slope downward ($\alpha < 0$), making
this parameter less relevant.  The range $-5.05 \leq \alpha \leq
-1.77$ encloses 90\% of the probability; the median value of $\alpha$
is -3.23.  The presence of the higher-mass samples in the analysis
produces a distinctive tail, eliminating the correlations discussed in
Section \ref{sec:power-law} and displayed in Figure
\ref{fig:power-law-2D} for the low-mass subset of the observations.

\begin{figure}
  \begin{center}
    \plotone{plots/power-law-high}
  \end{center}
  \caption{\label{fig:power-law-high} Histograms of the marginalized
    distribution for the three parameters $\Mmin$ (top, left), $\Mmax$
    (top, right), and $\alpha$ (bottom) from the power-law model
    including the higher-mass samples in the MCMC.  The distribution
    for $\Mmax$ is quite broad because the best fit power laws slope
    downward ($\alpha < 0$), making this parameter less relevant.  The
    range $-5.05 \leq \alpha \leq -1.77$ encloses 90\% of the
    probability; the median value of $\alpha$ is -3.23.  The presence
    of the higher-mass samples in the analysis produces a distinctive
    tail, eliminating the correlations discussed in Section
    \ref{sec:power-law} and displayed in Figure \ref{fig:power-law-2D}
    for the low-mass subset of the observations. }
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/exp-cutoff-high}
  \end{center}
  \caption{\label{fig:exp-cutoff-high} The marginalized distributions
    for the exponential parameters $\Mmin$ (top) and $M_0$ (bottom)
    defined in Section \ref{sec:exponential} from an analysis
    including the higher-mass systems.  The distribution for the scale
    mass, $M_0$, has moved to higher masses relative to Figure
    \ref{fig:exp-marginal} to fit the tail of the mass distribution;
    we now have $2.8292 \leq M_0 \leq 7.9298$ with 90\% confidence,
    with median 4.7003.  The distribution for $\Mmin$ is less
    affected, though it has broadened somewhat toward low masses.}
\end{figure}
